---
title: "HW 3 functions"
author: "Jasmine Syed"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

#Homework 1

##function 1

* calculates mean, variance, and std deviation of rnorm(10) and returns it as a list
```{r}

func1 <- function(x){
  a = sum(x)/length(x)
  b = sum((x-a)^2)/length(x)
  c = sqrt(b)
  return(list(mean=a,var=b,sd=c))
}

func1(rnorm(10))

```
##function 2

* calculates mean, variance, and std deviation of rnorm(10), and returns it as a list
* also, has checks that must pass in order for function to run properly
```{r}

func2 <- function(x){
  stopifnot(is.numeric(x))
  stopifnot(length(x)!=0)
  stopifnot(is.finite(x))
  stopifnot(!is.na(x))
  stopifnot(!is.nan(x))
  
  a = sum(x)/length(x)
  b = sum((x-a)^2)/length(x)
  c = sqrt(b)
  return(list(mean=a,var=b,sd=c))
}

func2(rnorm(10))

```
##function 3

* calculates mle using rnorm(10) as argument for function call
```{r warning=FALSE}

func3 <- function(x){
  alpha <- pi
  log <- function(alpha)
    sum(dgamma(x, shape = alpha, log = TRUE))
  interval <- mean(x) + c(-1,1) * 3 * sd(x)
  interval <- pmax(mean(x) / 1e3, interval)
  
  oout<- optimize(log, maximum = TRUE, interval)
  return (oout$maximum)
}

func3(rnorm(10))
```
##function 4

* using dataset d and its 2 factors, p (a probability) and x
* computing mean,variance, and std deviation of d and returning it as a list
```{r}

func4 <- function(d){
  
  a = sum(d$x * d$p)
  b = sum(((d$x - a)^2) * d$p)
  c = sqrt(b)
  return(list(mean=a,var=b,sd=c))
  
}
d <- read.table(url("http://www.stat.umn.edu/geyer/3701/data/q1p4.txt"),header = TRUE)
func4(d)
```

##function 5

* using dataset d and its 2 factors, p (a probability) and x
* goes through checks
```{r}

d <- read.table(url("http://www.stat.umn.edu/geyer/3701/data/q1p4.txt"),header = TRUE)
func5 <- function(d){
  
  stopifnot(is.numeric(d$x))
  stopifnot(is.numeric(d$p))
  
  stopifnot(length(d$x)!=0)
  stopifnot(length(d$p)!=0)
  
  stopifnot(is.finite(d$x))
  stopifnot(is.finite(d$p))
  
  stopifnot(!is.na(d$x))
  stopifnot(!is.na(d$p))
  
  stopifnot(!is.nan(d$x))
  stopifnot(!is.nan(d$p))
  
  stopifnot(all.equal(sum(d$p),1))
  
  a = sum(d$x * d$p)
  b = sum(((d$x - a)^2) * d$p)
  c = sqrt(b)
  return(list(mean=a,var=b,sd=c))
  
}
func5(d)
```

##function 6
* checks for problematic data, using NA in function call
```{r}


func6 <- function(x){
  
  tryCatch(stopifnot(is.numeric(x)), error=function(e){print("not numeric")})
  tryCatch(stopifnot(is.finite(x)), error=function(e){print("not finite")})
  tryCatch(stopifnot(length(x)!=0), error=function(e){print("has 0 length")})
  tryCatch(stopifnot(!is.nan(x)), error=function(e){print("NA or NAN")})
  tryCatch(stopifnot(!is.na(x)), error=function(e){print("NA or NAN")})
  
}

 func6(NA)

```

##function 7

* computing MLE for a gamma distribution
```{r warning=FALSE}
x <- scan(url("http://www.stat.umn.edu/geyer/3701/data/q1p3.txt"))
func7 <- function(x, func, interval){
  
  f7 <- function(theta, x)
  {sum(func(theta, x))}
  
  oout<- optimize(f7, maximum = TRUE, interval, x=x)
  return(oout$maximum)
} 
func1 = function(theta, x) dgamma(x, shape = theta, log = TRUE)
result7_gamma <- func7(x,func1,c(0,3))
result7_gamma

```


#Homework 2
##function 1

* matrix multiplication using t() and solve() and the matrix multiplication %*% 
```{r}


Fun1 <- function(A, x){
  load(url("http://www.stat.umn.edu/geyer/3701/data/q2p1.rda"))
  stopifnot(is.numeric(x))
  stopifnot(is.numeric(A))
  stopifnot(is.finite(x))
  stopifnot(is.finite(A))
  stopifnot(length(x) > 0)
  stopifnot(length(A) > 0)
  stopifnot(length(ncol(A)) > 0)
  stopifnot(length(nrow(A)) > 0)
  stopifnot(length(x)==nrow(A))  ##make sure length of x is equal to dimensions of square matrix 
  stopifnot(nrow(A) == ncol(A))  ##square matrix
  stopifnot(length(x)==ncol(A))
  
  
  
  t(x) %*% solve(A) %*% x    #matrix multiplication
}

load(url("http://www.stat.umn.edu/geyer/3701/data/q2p1.rda"))
Fun1(a, x)
```

##function 2

* matrix multiplication using t() and solve() and the matrix multiplication %*% 
* using binary operator as opposed to function call used in previous function
```{r}


"%Fun2%" <- function(A, x){
  stopifnot(is.numeric(x))
  stopifnot(is.numeric(A))
  stopifnot(is.finite(x))
  stopifnot(is.finite(A))
  stopifnot(length(x) > 0)
  stopifnot(length(A) > 0)
  stopifnot(length(ncol(A)) > 0)
  stopifnot(length(nrow(A)) > 0)
  stopifnot(length(x)==nrow(A))  ##make sure length of x is equal to dimensions of square matrix 
  stopifnot(nrow(A) == ncol(A))  ##square matrix
  stopifnot(length(x)==ncol(A))
  
  
  t(x) %*% solve(A) %*% x    #matrix multiplication
}
load(url("http://www.stat.umn.edu/geyer/3701/data/q2p1.rda"))
a %Fun2% x
```

##function 3 & 4

* standardizing each column of a matrix
```{r}

Fun3 <- function(A){
  load(url("http://www.stat.umn.edu/geyer/3701/data/q2p3.rda"))
  stopifnot(is.numeric(A))
  stopifnot(is.finite(A))
  stopifnot(length(A) > 0)
  stopifnot(ncol(A) > 0)
  stopifnot(nrow(A) > 1) 
  
  mean <- colMeans(A)
  sd <- apply(A,2,"sd")
  ( A - mean)/sd
  
}

load(url("http://www.stat.umn.edu/geyer/3701/data/q2p3.rda"))
Fun3(a)
```

##function 5

* examples for function comparable to apply()
```{r}

myapply <- function(X, MARGIN, FUN, ...)
{
  
  #stopifnot(length(dim(X))==2)
  
  if(length(dim(X))!=2)
  {
    stop("matrix is not 2d")
  } 
  if(!(MARGIN %in% c(1,2)))
  {
    stop("margin is not in 1 or 2")
  }
  R = dim(X)[1]
  C = dim(X)[2]
  f = match.fun(FUN)
  
  if (MARGIN == 1)
  {
    result = list()
    for(i in 1:R)
    {
      result[[i]] = f(X[i,],...)
    }
  }else if(MARGIN == 2)
  {
    result = list()
    for(j in 1:C)
    {
      result[[j]] = f(X[,j],...)
    }
  }
  return(simplify2array(result))
}

fred <- matrix(1:6, ncol = 2)
myapply(fred,1,"mean")
myapply(fred,2,"mean")

 myapply(fred,1,"max")
 myapply(fred,2,"max")
 
 myapply(fred,1,function(x) quantile(x,0.75))
 myapply(fred,2,function(x) quantile(x,0.75))
 
 myapply(fred, 1, quantile, probs=.75)
 myapply(fred, 2, quantile, probs=.75)
 
 myapply(fred, 1, quantile, probs = (1:3)/4)
 myapply(fred, 2, quantile, probs = (1:3)/4)
 
 moo <- myapply(fred, 1, mean)
 fred.mean.free.rows <- sweep(fred, 1, moo)
 rowMeans(fred.mean.free.rows)

```

#Additional functions

##ggplot function - dataset:d 
* relationship between x and each x's probability in the dataset d
```{r}

plotMyData<-function(x){
  library(magrittr)
 x%>% ggplot2::ggplot()+ggplot2::aes(x=x, y=p)+ggplot2::geom_point()
}
d <- read.table(url("http://www.stat.umn.edu/geyer/3701/data/q1p4.txt"),header = TRUE)
plotMyData(d)
```


##ggplot function - dataset:g

* since dataset used does not have two factors to study like dataset d, I made y variable a constant
```{r}
plotMyData<-function(x){
  library(magrittr)
 x%>% ggplot2::ggplot()+ggplot2::aes(x=x, y=0)+ggplot2::geom_point()
}

g <- scan(url("http://www.stat.umn.edu/geyer/3701/data/q1p1.txt"))
g<- as.data.frame(g)
plotMyData(g)
```

##ggplot function - dataset:b

* since dataset used does not have two factors to study like dataset d, I made y variable a function of x
```{r}

plotMyData<-function(x){
  library(magrittr)
 x%>% ggplot2::ggplot()+ggplot2::aes(x=x, y=1/x)+ggplot2::geom_point()
}

b <- scan(url("http://www.stat.umn.edu/geyer/3701/data/q1p3.txt"))
b<- as.data.frame(b)
plotMyData(b)
```
##dplyr function

* filters dataset d by showing only values with a p-value greater than 0.05
```{r}
filterFun<-function(x){
  dplyr::filter(x, p>0.05) 
}

d <- read.table(url("http://www.stat.umn.edu/geyer/3701/data/q1p4.txt"),header = TRUE)
filterFun(d)
```


